---
title: "Interactive Data Visualization in R with Shiny"
subtitle: "Day 1: Basic Shiny App - March 25th, 2026"
author: "Peter Lawson"
date: 2026-03-20
format:
  html:
    theme: cosmo
    toc: true
filters:
  - quarto
  - line-highlight
---

# Overview

Welcome to **Interactive Data Visualization in R with Shiny**. This is a step-by-step guide walking through the code we will write together during our hands-on session.  

  
## Hello World

Let's begin by building the simplest Shiny application we can. 

- Create minimum skeleton for a Shiny app. Name file `app.R` to have it recognized as a shiny application. 

- Add a title to the `page_sidebar` labeled `"Hello World!"`. 

```{r}
#| echo: true
#| eval: false
#| file: "src/part-1-introduction/01-hello-world.R"
```

In this step-by-step lesson, we will create a simple Shiny application. To see the application we will be building, you can run:

```{r}
#| echo: true
#| eval: false
library(shiny)
runExample("01_hello")
```

## Step 1: Skeleton
```{r}
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/02-old-faithful-skeleton.R"
```

## Step 2: Add Title
```{r}
#| source-line-numbers: "6-8"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/03-old-faithful-add-title.R"
```


## Step 3: Add Layout
```{r}
#| source-line-numbers: "9"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/04-old-faithful-add-layout.R"
```


## Step 4: Add Input
```{r}
#| source-line-numbers: "9-19"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/05-old-faithful-add-input.R"
```

## Step 5: Add Reactive Output
```{r}
#| source-line-numbers: "20-32"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/06-old-faithful-add-reactive-output.R"
```

## Step 6: Add Plot Output
```{r}
#| source-line-numbers: "15-17"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/07-old-faithful-add-plot-output.R"
```

## Step 7: Add Eruptions Histogram 
```{r}
#| source-line-numbers: "17-18,39-48"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/08-old-faithful-add-eruptions-histogram.R"
```
