---
title: "Interactive Data Visualization in R with Shiny"
subtitle: "Teaching Notebook - March 20th, 2026"
author: "Peter Lawson"
date: 2026-03-20
format:
  html:
    theme: cosmo
    toc: true
filters:
  - quarto
  - line-highlight
---

# Overview

Welcome to **Interactive Data Visualization in R with Shiny**. This is a step-by-step guide walking through the code we will write together during our hands-on session.  

## Learning Objectives

By the end of our two-part Shiny workshop, you should be able to: 

> 1. Articulate basic principles of interactive data visualization 
>
> 2. Understand critical concepts from R-Shiny including the structure of a Shiny App, how to run a Shiny App, how to configure the aesthetics of a Shiny App, and how the reactive programming model of Shiny works. 
>
> 3. Be able to augment a barebones Shiny app framework to develop a fully functioning Shiny app using provided data. This will encompass R and HTML programming. 
>
> 4. Be able to identify tools and resources for sharing and deploying a Shiny app. Be aware of advanced style configuration using CSS. 
    
  
## Hello World

Let's begin by building the simplest Shiny application we can. 

- Create minimum skeleton for a Shiny app. Name file `app.R` to have it recognized as a shiny application. 
- Add a title to the `page_sidebar` labeled `"Hello World!"`. 

```{r}
#| echo: true
#| eval: false
#| file: "src/part-1-introduction/01-hello-world.R"
```

# Part 1 - Developing a Simple Shiny Application

In this step-by-step lesson, we will create a simple Shiny application. To see the application we will be building, you can run:

```{r}
#| echo: true
#| eval: false
library(shiny)
runExample("01_hello")
```

## Step 1: Skeleton
```{r}
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/02-old-faithful-skeleton.R"
```

## Step 2: Add Title
```{r}
#| source-line-numbers: "6-8"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/03-old-faithful-add-title.R"
```

1. Add a title.

## Step 3: Add Layout
```{r}
#| source-line-numbers: "9"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/04-old-faithful-add-layout.R"
```


## Step 4: Add Input
```{r}
#| source-line-numbers: "9-19"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/05-old-faithful-add-input.R"
```

## Step 5: Add Reactive Output
```{r}
#| source-line-numbers: "20-32"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/06-old-faithful-add-reactive-output.R"
```

## Step 6: Add Plot Output
```{r}
#| source-line-numbers: "15-17"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/07-old-faithful-add-plot-output.R"
```

## Step 7: Add Eruptions Histogram 
```{r}
#| source-line-numbers: "16-17,39-48"
#| echo: true 
#| eval: false
#| file: "src/part-1-introduction/08-old-faithful-add-eruptions-histogram.R"
```

# Part 2 - Developing a Penguin Dashboard Application

## Step 1: Skeleton
```{r}
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/01-skeleton.R"
```

## Step 2: Data Import and Cleaning
```{r}
#| source-line-numbers: "3-22"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/02-data-import-cleaning.R"
```

## Step 3: Data Description
```{r}
#| source-line-numbers: "27-35"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/03-data-description.R"
```

## Step 4: Data Visualization
```{r}
#| source-line-numbers: "4,45-58"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/04-data-visualization.R"
```

## Step 5: Interactive Visualization
```{r}
#| source-line-numbers: "30-43,65-68"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/05-interactive-visualization.R"
```

## Step 6: Species Selector
```{r}
#| source-line-numbers: "42-51,70-73,78"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/06-species-selector.R"
```

## Step 7: Consistent Scale
```{r}
#| source-line-numbers: "84-90"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/07-consistent-scale.R"
```

## Step 8: Consistent Coloring
```{r}
#| source-line-numbers: "20-24,87-88"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/08-consistent-coloring.R"
```

## Step 9: Smoother and Facet
```{r}
#| source-line-numbers: "19,54-62,93-95"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/09-smoother-facet.R"
```

## Step 10: Add Summary
```{r}
#| source-line-numbers: "5,19-21,28,42,59-60,70-71,74,100-101,119-126"
#| echo: true 
#| eval: false
#| file: "src/part-2-penguins/10-add-summary.R"
```